\newpage
\section*{Lesson 1 Plan}

\textbf{Duration:} 60 minutes

\textbf{Lesson title:} Planning and Prototyping a Meaningful Minecraft Place

\textbf{Student-friendly target:} I can plan and begin building a Minecraft place that shows something important about identity or community, and I can explain how my design choices carry meaning.

\textbf{Identity-safe teacher language:} Your place can be real, imagined, symbolic, collaborative, or future-facing. You do not have to share anything private. The goal is to make thoughtful design choices and explain what those choices communicate.

\textbf{Criticality bridge:} \textcolor{LessonTwoAccent}{Students compare a generic Minecraft structure with a meaningful digital place and ask what makes a digital space readable, welcoming, accessible, or meaningful to an audience. They also begin asking whose stories are usually represented in digital spaces and whose stories are often missing.}

\textbf{Objectives:}
\begin{enumerate}[leftmargin=*]
    \item Students will choose a meaningful place connected to identity, community, belonging, or future possibility.
    \item Students will break the place into smaller build/code steps.
    \item Students will analyze how a sign, path, structure, color, repeated pattern, or coded feature can guide an audience's understanding.
    \item Students will begin questioning how digital spaces can include, exclude, preserve, or erase community stories.
\end{enumerate}

\renewcommand{\arraystretch}{0.82}
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\textwidth}{|>{\raggedright\arraybackslash}p{0.08\textwidth}|Y|Y|}
\hline
\rowcolor{LessonOneSoft}
\textbf{Mins} & \textbf{Activity} & \textbf{Purpose} \\ \hline
0--7 & Compare a generic Minecraft build with a meaningful digital place. Ask what makes a place meaningful, welcoming, or accessible. & Frames design as meaning-making and critical reading. \\ \hline
7--15 & Model a small place plan with entrance, sign, path, gathering spot, and repeated pattern. Think aloud about what each feature communicates and who the space welcomes. & Makes audience, access, and representation visible. \\ \hline
15--38 & Students choose a real, imagined, symbolic, or future place; sketch 3--5 features; and label first build/code steps. & Builds identity-safe agency and decomposition. \\ \hline
38--50 & Students begin a prototype, test one feature, and add one note explaining what the feature helps the audience understand. & Connects design to audience meaning. \\ \hline
50--60 & Partner share and exit slip: ``This helps my audience understand...'' and ``One story or access need I want to keep in mind is...'' & Captures oral, written, and critical evidence. \\ \hline
\end{tabularx}

\textbf{Differentiation:} Students may work alone or with roles such as builder, navigator, debugger, or explainer. Beginners may build one strong scene; advanced students may add loops, agent movement, signage, or an audience-centered accessibility feature.

\newpage
\section*{Lesson 2 Plan}

\textbf{Duration:} 60 minutes

\textbf{Lesson title:} Building, Coding, Debugging, and Analyzing the Meaningful Place

\textbf{Student-friendly target:} I can build and revise my Minecraft place so that at least one coded or repeated feature helps communicate meaning and support access.

\textbf{Joy + identity bridge:} Lesson 2 keeps coding joyful by treating debugging as experimentation rather than failure. Students revise features so their worlds become clearer, more welcoming, or more meaningful to an audience.

\textbf{Criticality bridge:} \textcolor{LessonTwoAccent}{Students think like designers and beginning developers. They do not only ask, ``Does it work?'' They also ask, ``Who does this design include? Who might it leave out? What could I change to make it more respectful, accessible, or clear?''}

\textbf{Objectives:}
\begin{enumerate}[leftmargin=*]
    \item Students will build at least two meaningful features from their plan.
    \item Students will use a sequence, loop, repeated structure, or agent-based action to create or improve the world.
    \item Students will debug one issue by naming the problem, testing a change, and recording what changed.
    \item Students will analyze how one code/build choice affects audience understanding, welcome, representation, or access.
\end{enumerate}

\renewcommand{\arraystretch}{0.82}
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\textwidth}{|>{\raggedright\arraybackslash}p{0.08\textwidth}|Y|Y|}
\hline
\rowcolor{LessonTwoSoft}
\textbf{Mins} & \textbf{Activity} & \textbf{Purpose} \\ \hline
0--8 & Students choose one must-build feature and one try-to-code feature, then name what the feature should help the audience understand. & Connects coding/design to meaning. \\ \hline
8--15 & Teacher models a debug cycle and how to write what the revision communicates or improves for access. & Makes debugging visible, writable, and purposeful. \\ \hline
15--42 & Students build and code with partner roles. Teacher asks: ``What does this choice help your audience read or understand? Who might need a clearer path, sign, explanation, or support?'' & Develops skill, analysis, and critical design. \\ \hline
42--52 & Screenshot checkpoint: students label one feature, explain how it works, explain what it communicates, and note one possible access or representation concern. & Connects product to intellectual and critical explanation. \\ \hline
52--60 & Quick share: one success, one bug, one next step, and one audience/access meaning. & Builds design language and community expertise. \\ \hline
\end{tabularx}

\textbf{Disciplinary writing evidence:} \textcolor{LessonTwoAccent}{Students complete the middle section of the Developer Design Note: ``My bug or design problem was...'' ``I expected...'' ``What happened was...'' ``I changed...'' ``This helped my audience understand...'' ``This made my world more welcoming/accessibile because...''}

\newpage
\section*{Lesson 3 Plan}

\textbf{Duration:} 60 minutes

\textbf{Lesson title:} Sharing Worlds, Reading Design, and Reflecting on Representation

\textbf{Student-friendly target:} I can present my Minecraft place, explain how it carries identity or community knowledge, and revise one feature so the world is clearer, more respectful, or more accessible.

\textbf{Representation bridge:} \textcolor{LessonTwoAccent}{Before students share, the teacher names that digital worlds are designed from choices. Some choices make people feel welcomed and seen; other choices erase people, flatten culture, or make access harder. Students explain what they chose, not private identity.}

\textbf{Objectives:}
\begin{enumerate}[leftmargin=*]
    \item Students will explain at least two design choices and how they carry meaning.
    \item Students will give peer feedback using notice/wonder language.
    \item Students will analyze how digital worlds can include, exclude, preserve, erase, flatten, or misrepresent people and communities.
    \item Students will revise one feature toward clarity, welcome, accessibility, respect, or repair.
\end{enumerate}

\renewcommand{\arraystretch}{0.82}
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\textwidth}{|>{\raggedright\arraybackslash}p{0.08\textwidth}|Y|Y|}
\hline
\rowcolor{LessonThreeSoft}
\textbf{Mins} & \textbf{Activity} & \textbf{Purpose} \\ \hline
0--8 & Students prepare a two-sentence explanation and choose one screenshot or live view to show. & Supports concise presentation. \\ \hline
8--35 & Gallery walk or paired screen-share. Students explain their worlds and collect one notice/wonder comment. & Centers oral, visual, and digital literacy. \\ \hline
35--48 & Whole-group critical discussion: ``Whose stories do digital worlds usually show, whose could be missing, who benefits from certain representations, and how can designers make spaces more respectful and accessible?'' & Moves Criticality from brainstorm into design reflection. \\ \hline
48--56 & Students revise one sentence, sign, pathway, structure, or accessibility feature based on feedback and explain why the change matters. & Turns feedback into critical action. \\ \hline
56--60 & Exit reflection: ``My world shows...'' ``A designer can make digital spaces more welcoming by...'' ``One feature I revised for access, respect, or repair was...'' & Captures final evidence. \\ \hline
\end{tabularx}

\textbf{Student reflection and analysis prompts:}
\begin{itemize}[leftmargin=*]
    \item \textcolor{LessonTwoAccent}{What does your Minecraft world help an audience understand?}
    \item \textcolor{LessonTwoAccent}{Which design choice carries the most meaning, and why?}
    \item \textcolor{LessonTwoAccent}{How do your paths, signs, colors, structures, repeated patterns, or code guide the audience?}
    \item \textcolor{LessonTwoAccent}{How can a digital world make someone feel welcomed or excluded?}
    \item \textcolor{LessonTwoAccent}{Whose stories are easy to represent in digital worlds, and whose stories are often missing?}
    \item \textcolor{LessonTwoAccent}{Who benefits when certain stories are centered, and who may be harmed or erased?}
    \item \textcolor{LessonTwoAccent}{How can designers avoid using culture, identity, or community knowledge as decoration?}
    \item \textcolor{LessonTwoAccent}{What did you revise after feedback, and how did that change your world's meaning, access, or respectfulness?}
\end{itemize}

\textbf{Developer Design Note prompt:} \textcolor{LessonTwoAccent}{``My design goal was...'' ``The build/code steps I used were...'' ``One bug or revision I worked through was...'' ``One feature I want my audience to notice is...'' ``This world communicates identity, community, belonging, or future possibility by...'' ``One revision I made for clarity, welcome, access, respect, or repair was...''}

\newpage
\section*{Assessment and Evidence}

\textbf{Evidence collected across the unit:} planning sheet, debug log, screenshots or sketches, teacher conference notes, peer feedback, short presentation, Developer Design Note, revision note, and exit reflection.

\renewcommand{\arraystretch}{0.95}
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\textwidth}{|>{\raggedright\arraybackslash}p{0.17\textwidth}|Y|Y|}
\hline
\rowcolor{SoftBlue}
\textbf{Criterion} & \textbf{Meeting} & \textbf{Developing} \\ \hline
Meaningful design & World includes features connected to identity, community, belonging, or future possibility. & World has features, but the meaning needs clearer explanation. \\ \hline
Computational thinking & Student decomposes, sequences, tests, debugs, or uses a repeated/coded feature. & Student builds but needs support naming steps or debugging. \\ \hline
Disciplinary writing & Student documents design goal, steps, bug/revision, and audience meaning using beginning computer science language. & Student explains the project generally but needs support connecting technical steps to meaning. \\ \hline
Intellectual analysis & Student analyzes how design choices such as paths, signs, structures, colors, repeated patterns, code, or accessibility supports help an audience read the world. & Student names a design choice but needs support explaining how it affects audience understanding. \\ \hline
Criticality & \textcolor{LessonTwoAccent}{Student questions representation, missing stories, access, and respectful design, then revises one feature toward clarity, welcome, access, respect, or repair.} & \textcolor{LessonTwoAccent}{Student notices representation or access with prompting but needs support connecting the concern to a design revision.} \\ \hline
\end{tabularx}

\textcolor{LessonTwoAccent}{\textbf{How the evidence is read:} The teacher does not grade cultural disclosure or artistic polish. The main evidence is whether students connect a design choice to a purpose, use basic computer science language to describe process, analyze how choices affect audience meaning, question representation/access, and revise after testing or feedback.}

\section*{Sample and Appendix Tools}

\textbf{Teacher model sample:} A small community garden world with an entrance path, welcome sign, gathering circle, repeated fence pattern, one coded/agent-assisted repeated feature, and one deliberately revised accessibility feature. The model should be meaningful but not perfect so students see revision as normal.

\textbf{Student-facing routine:} \textcolor{LessonTwoAccent}{Students use one tool at a time: plan first, debug next, then complete the Developer Design Note and share card. This keeps the workflow accessible while preserving a full evidence trail.}

\newpage
\section*{Reflection}

This mini-unit shows how digital literacy can become a space for belonging. Joy is built into choice, creation, debugging, peer support, and visible progress. Identity is built into meaningful place-making and student control over what is shared. Skills are built through sequencing, decomposition, debugging, documentation, and presentation. Intellect is built when students analyze how technical and design choices communicate meaning to an audience. \textcolor{LessonTwoAccent}{Criticality is built when students question representation, access, missing stories, and the power of digital design to include, exclude, preserve, erase, or repair.}

\textcolor{LessonTwoAccent}{The revision strengthens the mini-unit by making criticality visible and actionable. Students are not only building in Minecraft; they are learning how designers and beginning developers communicate a design goal, document a process, test and debug, respond to feedback, and explain how technical choices affect an audience. They also learn that code and digital space are not neutral. Code, signs, paths, colors, structures, repeated patterns, and accessibility supports can shape who feels welcomed, whose stories are seen, and whose knowledge is respected.}

\textcolor{LessonTwoAccent}{For Ethan specifically, the design matters because it does not ask him to simplify himself into one identity label. It gives him a way to connect coding, game design, land, community, mixed identity, and cultural knowledge only to the extent he chooses. The criticality layer matters because it protects that choice while still asking students to think honestly about representation, access, and respectful design.}

For my own teaching, the most important design decision is that students are assessed on how intentionally they communicate meaning and how thoughtfully they revise for audience, access, and respect. That distinction matters for neurodivergent, culturally connected, multilingual, and mixed-identity students because it keeps literacy tied to agency. In this unit, joy is not separate from rigor: students still decompose, debug, revise, document, analyze, and explain, but those skills serve a world they care about.

\textcolor{LessonTwoAccent}{This pursuit expands students' criticality because it helps them understand that design is a form of power. Students learn to read digital spaces as texts and to understand that technical choices are not neutral. A design can welcome, confuse, include, exclude, preserve, flatten, erase, or repair meaning. The work prepares students to see themselves not just as users of technology, but as designers whose choices affect others and whose revisions can move technology toward care, access, and transformation.}

\newpage
